Opening up the court (surface) in tennis grand slams

### 2018-09-15 contest/surface-type/paper-tennis.Rmd
merge

---
title: "Analyzing the Effect of Surface in Tennis Grand Slams"
author: "###########"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: true
header-includes:
  - \usepackage{hyperref}
abstract: "Tennis grand slams consist of the Australian Open, French Open, Wimbledon, and US Open, which are played on hard (Plexicushion), clay, grass, and hard (DecoTurf) courts, respectively.  The surface type may substantially impact ball speed, height, and spin as well as player speed and agility.  It is also believed that play style and practice habits may contribute to different results across surface types.  For example, Rafael Nadal is thought to be the best clay court player of all time whereas Roger Federer is particularly known for dominance at Wimbledon.  On the women's side, Serena Williams once struggled on clay courts but has seemingly transformed her style to perform better on clay courts, but has perhaps suffered on grass as a consequence.  In this analysis, we examine the result of the top 100 players in grand slams from 2013-2017 across the four different surfaces.  We create a hierarchical model with fixed and random effects to predict the number of points won in a match.  We take into consideration player-specific effects, nationality (which is thought to have an effect on play style), sex, ranking, ELO, and game statistics.   We assess the fit of our model using standard statistical techniques (e.g. MSE, AIC, BIC, residual diagnostics) in addition to `common knowledge' factors (for instance, Rafael Nadal should be indicated as a superior clay court player by the model).  We compare the results of top 100 players across grand slams to examine the effect of court surface. We also provide an in-depth analysis of Nadal, Federer, and S. Williams."
bibliography: master.bib
---

@@

```{r tab-data, include = TRUE, echo = FALSE}
w_id <- which(gs$winner_name == "Serena Williams")[1:2]
f_id <- which(gs$winner_name == "Roger Federer")[1:2]
n_id <- which(gs$winner_name == "Rafael Nadal")[1:2]
s_id <- which(gs$winner_name %in% c("Serena Williams", "Roger Federer", "Rafael Nadal"))
cols <- c("winner_name", "tourney_name", "year", "winner_ioc", "w_pointswon", "winner_rank", "l_pointswon", "loser_rank")
knitr::kable(gs[c(w_id, f_id, n_id), cols], format = "latex",
             row.names = FALSE, col.names = c("Winner", "Tournament", "Year", "W. IOC", "W. Points",  "W. Rank", "L. Points", "L. Rank"),
             caption = "\\label{tab:data}Example of the grand slam data.  It includes winner and loser attributes, match attributes, and tournament attributes.  Not all attributes are shown here.",
             booktabs = TRUE) %>% kable_styling(latex_options = "striped")
```

@@

## Examining individual players
```{r table}
data(gs_partial_players)

@@

knitr::kable(tab, format = "latex",
               row.names = FALSE, col.names = c("Tournament", "Nadal", "Federer", "Williams"),
               caption = "\\label{tab:three-gs-counts}Number of matches played for Nadal, Federer, and Williams from 2013-2017 at each of the grand slams.",
               booktabs = TRUE) %>% kable_styling(latex_options = "striped") %>%
    row_spec(5, bold = TRUE, background = "lightgray")
```

@@

We fit individual models for these three individuals (subsetting the data to their matches only), estimating the percent of points won in a match using stepwise regression.  The lower model is the  percent of points regressed on the opponent rank and indicator variables for court type with the FO as the reference variable.  The upper model is the lower model with the additional variables of winner to unforced error ratio, average serve speed, percent aces, percent break points, percent net points won, and their interaction effects with court type.  We do not display the full best fit models here, but they are available online.

ADD CIS!!

## Nadal

@@

coefs_n <- coefficients(step_model_n)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids} (along with the other linear regression diagnostics plots not shown here) shows that the model is a good fit for predicting the percent of points won.  For the significant variables ($\alpha$ = 95\%), while adjusting for the other variables, we expect percent of points won to be $`r signif(abs(coefs_n[2]), 2)`$ less at the AO and $`r signif(abs(coefs_n[4]), 2)`$ less at Wim compared to the FO; to increase by $`r signif(abs(coefs_n[5]), 2)`$ for a one unit increase in opponent rank; a `r .01 * signif(abs(coefs_n[7]), 2)`\% increase for a .01 increase in percent of break points won; and to decrease by $`r .01 * signif(abs(coefs_n[8]), 2)`$ for a .01 increase in percent of net points won.  As such, we see clear evidence Nadal performs better at the FO compared to the other grand slams.

## Federer
```{r federer-model}

@@

coefs_f <- coefficients(step_model_f)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}  shows that the model is a fair fit for predicting the percent of points won, but almost seem to see two clusters of residuals splitting with predicted point percentage of about 0.63.  For the significant variabless ($\alpha$ = 95\%), while adjusting for the other variables, we expect percent of points won to be $`r signif(abs(coefs_f[2] + coefs_f[8]), 2)`$ less at AO compared to the Wim.; $`r signif(abs(coefs_f[3] + coefs_f[9]), 2)`$ less at FO compared to the Wim.; $`r signif(abs(coefs_f[4] + coefs_f[9]), 2)`$\% less at USO compared to the Wim.;  to decrease by $`r signif(abs(coefs_f[6] + coefs_f[8]) , 2)`$ for a one unit increase in W/UE at AO, $`r signif(abs(coefs_f[6] + coefs_f[9]) , 2)`$ for a one unit increase in W/UE at FO, and $`r signif(abs(coefs_f[6] + coefs_f[10]) , 2)`$ for a one unit increase in W/UE at USO.  It is clear that Federer performs best at Wimbledon.

## Williams
```{r williams-model}

@@

coefs_w <- coefficients(step_model_w)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}  shows that the model is a fair fit for predicting the percent of points won, but we tend to overestimate the percent of points won in comparison with Nadal.   For the significant variables ($\alpha$ = 95\%), while adjusting for the other variables, we expect percent of points won to increase by $`r signif(abs(coefs_w[5]), 2)`$ for a 1 unit increase in oponent rank; to increase by $`r signif(abs(coefs_w[6] + coefs_w[11]) , 2)`$ for a one unit increase in W/UE at FO, to increase by  $`r signif(abs(coefs_w[6] + coefs_w[12]) , 2)`$for a one unit increase in W/UE at USO, and to increase by  $`r signif(abs(coefs_w[6] + coefs_w[13]) , 2)`$ for a one unit increase in W/UE at Wim. compared to AO; to increase by  $`r signif(.01 * abs(coefs_w[8] + coefs_w[17]) , 2)`$ for a .01 increase in percent of aces at the French Open; and to increase by $`r signif(.01 * abs(coefs_w[9] + coefs_w[20]) , 2)`$ for a .01 increase in percent of break points won at the French Open.

>>>>

---
title: "Opening up the court (surface) in tennis grand slams"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: true
header-includes:
  - \usepackage{hyperref}
abstract: "Tennis grand slams consist of the Australian Open, French Open, Wimbledon, and US Open, which are played on hard (Plexicushion), clay, grass, and hard (DecoTurf) courts, respectively.  The surface type may substantially impact ball speed, height, and spin as well as player speed and agility.  It is also believed that play style and practice habits may contribute to different results across surface types.  For example, Rafael Nadal is thought to be the best clay court player of all time whereas Roger Federer is particularly known for dominance at Wimbledon.  On the women's side, Serena Williams once struggled on clay courts but has seemingly transformed her style to perform better on clay courts, but has perhaps suffered on grass as a consequence.  In this analysis, we examine the result of the top 100 players in grand slams from 2013-2017 across the four different surfaces.  We create a hierarchical model with fixed and random effects to predict the number of points won in a match.  We take into consideration player-specific effects, nationality (which is thought to have an effect on play style), sex, ranking, ELO, and game statistics.   We assess the fit of our model using standard statistical techniques (e.g. MSE, AIC, BIC, residual diagnostics) in addition to `common knowledge' factors (for instance, Rafael Nadal should be indicated as a superior clay court player by the model).  We compare the results of top 100 players across grand slams to examine the effect of court surface. We also provide an in-depth analysis of Nadal, Federer, and S. Williams."
bibliography: master.bib
---

@@

```{r tab-data, include = TRUE, echo = FALSE}
w_id <- which(gs$winner_name == "Serena Williams")[1:2]
f_id <- which(gs$winner_name == "Roger Federer")[1:2]
n_id <- which(gs$winner_name == "Rafael Nadal")[1:2]
s_id <- which(gs$winner_name %in% c("Serena Williams", "Roger Federer", "Rafael Nadal"))
cols <- c("winner_name", "tourney_name", "year", "winner_ioc", "w_pointswon", "winner_rank", "l_pointswon", "loser_rank")
knitr::kable(gs[c(w_id, f_id, n_id), cols], format = "latex",
             row.names = FALSE, col.names = c("Winner", "Tournament", "Year", "W. IOC", "W. Points",  "W. Rank", "L. Points", "L. Rank"),
             caption = "\\label{tab:data}Example of the grand slam data.  It includes winner and loser attributes, match attributes, and tournament attributes.  Not all attributes are shown here.",
             booktabs = TRUE) %>% kable_styling(latex_options = "striped", "hold_position")
```

@@

## Examining individual players

```{r table, fig.pos = 'H'}
data(gs_partial_players)

@@

knitr::kable(tab, format = "latex",
               row.names = FALSE, col.names = c("Tournament", "Nadal", "Federer", "Williams"),
               caption = "\\label{tab:three-gs-counts}Number of matches played for Nadal, Federer, and Williams from 2013-2017 at each of the grand slams.",
               booktabs = TRUE) %>% kable_styling(latex_options = "striped", "HOLD_position") %>%
    row_spec(5, bold = TRUE, background = "lightgray")
```

@@

We fit individual linear models for these three individuals (subsetting the data to their matches only), estimating the percent of points won in a match using stepwise regression.  The lower model is the  percent of points regressed on the opponent rank and indicator variables for court type with the FO as the reference variable.  The upper model is the lower model with the additional variables of winner to unforced error ratio (W/UE), average serve speed, percent aces, percent break points, percent net points won, and their interaction effects with court type.  We do not display the full best fit models here, but they are available online.  We do report the effects ($\pm 2 \cdot$ Std. Err.) of significant variables and their interaction terms, adjusting for the other covariates.  Since we postulated that Nadal is best on clay, Federer on grass, and Williams on hard court, specificically, AO, we use the FO, Wim., and AO, respectively, as the reference variable in the model.  For Nadal and Federer we use forward-backwards stepwise regression, using AIC as the criterion, beginning with the full model.  For Williams, we also use forwards-backwards stepwise regression but begin at the low model since she has more missing data.

## Nadal: King of Clay: confirmed

@@

coefs_n <- coefficients(step_model_n)
cov_n <- vcov(step_model_n)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids} (along with the other linear regression diagnostics plots not shown here) shows that the model is a good fit for predicting the percent of points won.  For the significant variables ($\alpha$ = 95\%), while adjusting for the other variables, we expect percent of points won to be $`r signif(abs(coefs_n[2]), 2)` \pm `r signif(2 * sqrt(cov_n[2,2]), 2)`$ less at the AO and $`r signif(abs(coefs_n[4]), 2)` \pm `r signif(2 * sqrt(cov_n[4,4]), 2)`$ less at Wim compared to the FO; to increase by $`r signif(abs(coefs_n[5]), 2)`\pm `r signif(2 * sqrt(cov_n[5,5]), 2)`$ for a one unit increase in opponent rank; a $`r .01 * signif(abs(coefs_n[7]), 2)`\pm `r signif(.02 * sqrt(cov_n[7,7]), 2)`$ increase for a .01 increase in percent of break points won; and to decrease by $`r .01 * signif(abs(coefs_n[8]), 2)` \pm `r signif(.02 * sqrt(cov_n[8,8]), 2)`$ for a .01 increase in percent of net points won.  As such, we see clear evidence Nadal performs better at the FO compared to the other grand slams.

## Federer: Wimbledon extraordinaire: debunked
```{r federer-model}

@@

coefs_f <- coefficients(step_model_f)
cov_f <- vcov(step_model_f)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}  shows that the model is a fairly good fit for predicting the percent of points won, but almost seem to see two clusters of residuals splitting with predicted point percentage of about 0.63.  For the significant variables ($\alpha$ = 95\%), while adjusting for the other variables, we expect percent of points won to be $`r signif(abs(coefs_f[2] + coefs_f[8]), 2)` \pm `r signif(2 * sqrt(cov_f[2,2] + cov_f[8,8] + 2 * cov_f[2,8]), 2)`$ less at AO compared to the Wim.; $`r signif(abs(coefs_f[3] + coefs_f[9]), 2)`\pm `r signif(2 * sqrt(cov_f[3,3] + cov_f[9,9] + 2 * cov_f[3,9]), 2)`$ less at FO compared to the Wim.; $`r signif(abs(coefs_f[4] + coefs_f[9]), 2)`\pm `r signif(2 * sqrt(cov_f[4,4] + cov_f[9,9] + 2 * cov_f[4,9]), 2)`$ less at USO compared to the Wim.;  to decrease by $`r signif(abs(coefs_f[6] + coefs_f[8]) , 2)` \pm `r signif(2 * sqrt(cov_f[6,6] + cov_f[8,8] + 2 * cov_f[6,8]), 2)`$ for a one unit increase in W/UE at AO, $`r signif(abs(coefs_f[6] + coefs_f[9]) , 2)` \pm `r signif(2 * sqrt(cov_f[6,6] + cov_f[9,9] + 2 * cov_f[6,9]), 2)`$ for a one unit increase in W/UE at FO, and $`r signif(abs(coefs_f[6] + coefs_f[10]) , 2)` \pm `r signif(2 * sqrt(cov_f[6,6] + cov_f[10,10] + 2 * cov_f[6,10]), 2)`$ for a one unit increase in W/UE at USO.  This is strong evidence that Federer performs best at Wimbledon but possibly does better at the USO, depending on his W/UE ratio.

## Williams: hard court magician only: debunked
```{r williams-model}

@@

cov_w <- vcov(step_model_w)
coefs_w <- coefficients(step_model_w)
```

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}  shows that the model is a fair fit for predicting the percent of points won, but we tend to overestimate the percent of points won in comparison with Nadal.   For the significant variables ($\alpha$ = 95$\%$), while adjusting for the other variables, we expect percent of points won to increase by $`r signif(abs(coefs_w[5]), 2)` \pm `r signif(2 * sqrt(cov_w[5,5]), 2)`$ for a 1 unit increase in oponent rank; to increase by $`r signif(abs(coefs_w[6] + coefs_w[11]) , 2)` \pm `r signif(2 * sqrt(cov_w[6,6] + cov_w[11,11] + 2 * cov_w[6,11]), 2)`$ for a one unit increase in W/UE at FO, to increase by  $`r signif(abs(coefs_w[6] + coefs_w[12]) , 2)` \pm `r signif(2 * sqrt(cov_w[6,6] + cov_w[12,12] + 2 * cov_w[6,12]), 2)`$ for a one unit increase in W/UE at USO, and to increase by  $`r signif(abs(coefs_w[6] + coefs_w[13]) , 2) `\pm `r signif(2 * sqrt(cov_w[6,6] + cov_w[13,13] + 2 * cov_w[6,13]), 2)`$ for a one unit increase in W/UE at Wim. compared to AO; to increase by  $`r signif(.01 * abs(coefs_w[8] + coefs_w[17]) , 2)`\pm `r signif(.02 * sqrt(cov_w[8,8] + cov_w[17,17] + 2 * cov_w[17,8]), 2)`$ for a .01 increase in percent of aces at the French Open; and to increase by $`r signif(.01 * abs(coefs_w[9] + coefs_w[20]) , 2)` \pm `r signif(.02 * sqrt(cov_w[9,9] + cov_w[20,20] + 2 * cov_w[20,9]), 2)`$ for a .01 increase in percent of break points won at the French Open.

### 2018-09-15 contest/surface-type/paper-tennis.tex
merge

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Analyzing the Effect of Surface in Tennis Grand Slams},
            pdfauthor={Kayla Frisoli, Shannon Gallagher, and Amanda Luby},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Analyzing the Effect of Surface in Tennis Grand Slams}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Kayla Frisoli, Shannon Gallagher, and Amanda Luby}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{September 15, 2018}

@@

\hypertarget{sec:iintro}{%
\section{Introduction}\label{sec:iintro}}

@@

\includegraphics{paper-tennis_files/figure-latex/unnamed-chunk-2-1.pdf}

@@

\includegraphics{paper-tennis_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

We fit individual models for these three individuals (subsetting the
data to their matches only), estimating the percent of points won in a
match using stepwise regression. The lower model is the percent of
points regressed on the opponent rank and indicator variables for court
type with the FO as the reference variable. The upper model is the lower
model with the additional variables of winner to unforced error ratio,
average serve speed, percent aces, percent break points, percent net
points won, and their interaction effects with court type. We do not
display the full best fit models here, but they are available online.

ADD CIS!!

\hypertarget{nadal}{%
\subsection{Nadal}\label{nadal}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
(along with the other linear regression diagnostics plots not shown
here) shows that the model is a good fit for predicting the percent of
points won. For the significant variables (\(\alpha\) = 95\%), while
adjusting for the other variables, we expect percent of points won to be
\(0.034\) less at the AO and \(0.046\) less at Wim compared to the FO;
to increase by \(1.5\times 10^{-4}\) for a one unit increase in opponent
rank; a 0.0012\% increase for a .01 increase in percent of break points
won; and to decrease by \(8.8\times 10^{-4}\) for a .01 increase in
percent of net points won. As such, we see clear evidence Nadal performs
better at the FO compared to the other grand slams.

\hypertarget{federer}{%
\subsection{Federer}\label{federer}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
shows that the model is a fair fit for predicting the percent of points
won, but almost seem to see two clusters of residuals splitting with
predicted point percentage of about 0.63. For the significant variabless
(\(\alpha\) = 95\%), while adjusting for the other variables, we expect
percent of points won to be \(0.04\) less at AO compared to the Wim.;
\(0.058\) less at FO compared to the Wim.; \(0.007\)\% less at USO
compared to the Wim.; to decrease by \(0.022\) for a one unit increase
in W/UE at AO, \(0.073\) for a one unit increase in W/UE at FO, and
\(0.041\) for a one unit increase in W/UE at USO. It is clear that
Federer performs best at Wimbledon.

\hypertarget{williams}{%
\subsection{Williams}\label{williams}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
shows that the model is a fair fit for predicting the percent of points
won, but we tend to overestimate the percent of points won in comparison
with Nadal. For the significant variables (\(\alpha\) = 95\%), while
adjusting for the other variables, we expect percent of points won to
increase by \(4.8\times 10^{-4}\) for a 1 unit increase in oponent rank;
to increase by \(0.034\) for a one unit increase in W/UE at FO, to
increase by \(0.092\)for a one unit increase in W/UE at USO, and to
increase by \(0.0038\) for a one unit increase in W/UE at Wim. compared
to AO; to increase by \(2\times 10^{-4}\) for a .01 increase in percent
of aces at the French Open; and to increase by \(0.0024\) for a .01
increase in percent of break points won at the French Open.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Opening up the court (surface) in tennis grand slams},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Opening up the court (surface) in tennis grand slams}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{September 15, 2018}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{sec:iintro}{%
\section{Introduction}\label{sec:iintro}}

@@

\begin{center}\includegraphics{paper-tennis_files/figure-latex/unnamed-chunk-2-1} \end{center}

@@

\begin{center}\includegraphics{paper-tennis_files/figure-latex/unnamed-chunk-3-1} \end{center}

\hypertarget{tall-players-and-aces}{%
\subsection{Tall players and aces}\label{tall-players-and-aces}}

Why the actual EFF won't fig.width work?

\begin{center}\includegraphics{paper-tennis_files/figure-latex/unnamed-chunk-4-1} \end{center}

@@

We fit individual linear models for these three individuals (subsetting
the data to their matches only), estimating the percent of points won in
a match using stepwise regression. The lower model is the percent of
points regressed on the opponent rank and indicator variables for court
type with the FO as the reference variable. The upper model is the lower
model with the additional variables of winner to unforced error ratio
(W/UE), average serve speed, percent aces, percent break points, percent
net points won, and their interaction effects with court type. We do not
display the full best fit models here, but they are available online. We
do report the effects (\(\pm 2 \cdot\) Std. Err.) of significant
variables and their interaction terms, adjusting for the other
covariates. Since we postulated that Nadal is best on clay, Federer on
grass, and Williams on hard court, specificically, AO, we use the FO,
Wim., and AO, respectively, as the reference variable in the model. For
Nadal and Federer we use forward-backwards stepwise regression, using
AIC as the criterion, beginning with the full model. For Williams, we
also use forwards-backwards stepwise regression but begin at the low
model since she has more missing data.

\hypertarget{nadal-king-of-clay-confirmed}{%
\subsection{Nadal: King of Clay:
confirmed}\label{nadal-king-of-clay-confirmed}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
(along with the other linear regression diagnostics plots not shown
here) shows that the model is a good fit for predicting the percent of
points won. For the significant variables (\(\alpha\) = 95\%), while
adjusting for the other variables, we expect percent of points won to be
\(0.034 \pm 0.028\) less at the AO and \(0.046 \pm 0.033\) less at Wim
compared to the FO; to increase by
\(1.5\times 10^{-4}\pm 1.3\times 10^{-4}\) for a one unit increase in
opponent rank; a \(0.0012\pm 5.6\times 10^{-4}\) increase for a .01
increase in percent of break points won; and to decrease by
\(8.8\times 10^{-4} \pm 5.5\times 10^{-4}\) for a .01 increase in
percent of net points won. As such, we see clear evidence Nadal performs
better at the FO compared to the other grand slams.

\hypertarget{federer-wimbledon-extraordinaire-debunked}{%
\subsection{Federer: Wimbledon extraordinaire:
debunked}\label{federer-wimbledon-extraordinaire-debunked}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
shows that the model is a fairly good fit for predicting the percent of
points won, but almost seem to see two clusters of residuals splitting
with predicted point percentage of about 0.63. For the significant
variables (\(\alpha\) = 95\%), while adjusting for the other variables,
we expect percent of points won to be \(0.04 \pm 0.034\) less at AO
compared to the Wim.; \(0.058\pm 0.046\) less at FO compared to the
Wim.; \(0.007\pm 0.093\) less at USO compared to the Wim.; to decrease
by \(0.022 \pm 0.019\) for a one unit increase in W/UE at AO,
\(0.073 \pm 0.066\) for a one unit increase in W/UE at FO, and
\(0.041 \pm 0.033\) for a one unit increase in W/UE at USO. This is
strong evidence that Federer performs best at Wimbledon but possibly
does better at the USO, depending on his W/UE ratio.

\hypertarget{williams-hard-court-magician-only-debunked}{%
\subsection{Williams: hard court magician only:
debunked}\label{williams-hard-court-magician-only-debunked}}

@@

The standardized residuals plot shown in Figure \ref{fig:individ-resids}
shows that the model is a fair fit for predicting the percent of points
won, but we tend to overestimate the percent of points won in comparison
with Nadal. For the significant variables (\(\alpha\) = 95\(\%\)), while
adjusting for the other variables, we expect percent of points won to
increase by \(4.8\times 10^{-4} \pm 3.2\times 10^{-4}\) for a 1 unit
increase in oponent rank; to increase by \(0.034 \pm 0.02\) for a one
unit increase in W/UE at FO, to increase by \(0.092 \pm 0.045\) for a
one unit increase in W/UE at USO, and to increase by \(0.0038\pm 0.066\)
for a one unit increase in W/UE at Wim. compared to AO; to increase by
\(2\times 10^{-4}\pm 1.6\times 10^{-4}\) for a .01 increase in percent
of aces at the French Open; and to increase by \(0.0024 \pm 0.0011\) for
a .01 increase in percent of break points won at the French Open.
